Bioconductor workflow for single-cell RNA sequencing: Normalization, dimensionality reduction, clustering, and lineage inference

### 2017-06-30 workflow/workflow.Rmd
Updates to text.
change to workflow

EAP: small request. Can everyone put a line between the beginning of a r chunk and text? It makes it nicely formated for my text editor.

@@

Overall, after the pre-processing steps, our dataset has 1,000 genes and 747 cells.
```{r}
core
```

@@

The next step of the workflow is to cluster the cells according to the low-dimensional matrix `W` computed in the previous step. We use the resampling-based sequential ensemble clustering (RSEC) frameworkd implemented in the `RSEC` function from the Bioconductor R package `clusterExperiment`. Specifically, given a set of user-supplied based clustering algorithms and associated tuning parameters (e.g., _k_-means, with a range of values for _k_), RSEC generates a collection of candidate clusterings by resampling cells and using a sequential tight clustering procedure as in Tseng and Wong (2005)[@Tseng2005]. A consensus clustering is obtained based on the co-clustering matrix and non-differential clusters are then merged by creating a hierarchy of clusters, working up the tree, testing for differential expression between sister nodes, and collapsing nodes with insufficient DE. As in supervised learning, resampling greatly improves the stability of clusters and considering an ensemble of methods and tuning parameters allows us to capitalize on the different strengths of the base algorithms and avoid the subjective selection of tuning parameters.

Elizabeth, would you like to more details on the rational and arguments of the different functions?

@@

@@

@@

More formally, we are interested in testing for variable expression over pseudotime. A general additive model (GAM) is a flexible, nonlinear way to model the relationship between a gene's normalized expression measures and pseudotime. We implemented this model for each gene and show the top 100 genes by p-value in the heatmap below.

@@

>>>>

@@

Overall, after the pre-processing steps, our dataset has 1,000 genes and 747 cells.

```{r}
core
```

@@

The next step of the workflow is to cluster the cells according to the low-dimensional matrix `W` computed in the previous step. We use the resampling-based sequential ensemble clustering (RSEC) frameworkd implemented in the `RSEC` function from the Bioconductor R package `clusterExperiment`. Specifically, given a set of user-supplied based clustering algorithms and associated tuning parameters (e.g., _k_-means, with a range of values for _k_), RSEC generates a collection of candidate clusterings by resampling cells and using a sequential tight clustering procedure as in Tseng and Wong (2005)[@Tseng2005]. A consensus clustering is obtained based on the levels of co-clustering of samples across the candidate clusterings. The consensus clustering is further condensed by merging similar clusters which is done by creating a hierarchy of clusters, working up the tree, and testing for differential expression between sister nodes, with nodes of insufficient DE collapsed. As in supervised learning, resampling greatly improves the stability of clusters and considering an ensemble of methods and tuning parameters allows us to capitalize on the different strengths of the base algorithms and avoid the subjective selection of tuning parameters.

@@

EP: the label `clusterLabels` for the original labels might do better to be changed to `publishedClusters` so that it is more informative. Also, we don't do a direct comparison of the `publishedClusters`
 to the found clusters -- is that relevant? If so, it is probably better to add the published clusters via `addClusters` so it is easier to treat them like clusters. This is perhaps useful for a workflow since this is a common task that `clusterExperiment` makes easy. Here is code that does this:

 ```{r addPublishedClusters}
 #delete this chunk if not of interest
 #if keep, need to choose if do both figures and add figure caption 
 ceObj<-addClusters(ceObj,colData(ceObj)$clusterLabels,clusterLabel="publishedClusters")
 plotClusters(ceObj, whichClusters=c("mergeClusters","combineMany","publishedClusters"))
 plotBarplot(ceObj,whichClusters=c("mergeClusters","publishedClusters")) 
 ```

 EP: If we do the above code, then the `plotHeatmap` command below should be modified to have the `publishedClusters` under the `whichClusters` argument.

@@

[SD: Mention reason for having these two figures. What do colors correspond to in second figure? Match notation for dimensions (along diagonal) in the two figures.]

@@

More formally, we are interested in testing for variable expression over pseudotime. We do this by fitting a loess curve to the data which is a flexible, nonlinear way to model the relationship between a gene's normalized expression measures and pseudotime. We then can test the null hypothesis of no change over time using the `gam` package. We implemented this model for each gene and show the top 100 genes by p-value in the heatmap below.

@@

[SD: Add figure caption.]
